More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-03-28 2016-Weak-FH-Bisim/WeakBisim.tex
algo, elim, WOT

\TODO{Eric: refine the algorithm description}

@@

2) While there are unexplored states:

2a) pick one state from the unexplored set and add it to the explored
set. From this state
build all possible deduction trees by application of the structural
rules Tr1 and Tr2, using all \ERIC{applicable} combinations
of synchronisation vectors.

2b) For each of the obtained deduction trees:\\
- Extract the resulting open-transition, with its predicate and
\Post\ assignments by exploring the structure of the pNet.

- The predicate is submitted to Z3 for checking satisfiability. If it
is NOT stisfiable, the resulting OT is discarded. This will minimize
the number of resulting transitions, and potentialy prune the search-space.

- For each open-transition,  add the transition in the outgoing transitions of
the current state, and add the
resulting state in the unexplored set if it is not already there.

%3) this ends with the "while" from step 2, when we have no more
%unexplored set.
\end{algorithm}

Given an open-pNet
with finite synchronisation sets, finitely many leaves and
holes, and each pLTS at leaves having a finite number of states and
(symbolic) transitions, has a finite automaton (see \ref{Forte}).

@@

\ERIC{Optionally, simplify the predicate by eliminating the unnecessary intermediate variables, that were produced as fresh local variables of synchronisation vectors, and fresh result variables of intermediate OTs. In the resulting predicate the only significant variables are :\\
- the input variables of pLTS transitions\\
- the actions of holes\\
- the result action at toplevel.}

  \TODO{Eric: define ``elimination of intermediate variables ?'' it
    transforms the predicates, it is not an
    ``equivalence'' transformation, rather some sort of
    abstraction... Does it change the satisfiability ?}

%% \begin{theorem}[Finiteness of open-automata.]\\
%% Given an open pNet $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in K}\myrangle$ with leaves $pLTS_i^{i\in L}$ and holes $Hole_j^{j\in
%%   J}$, if the sets $L$ and $J$ are finite, if the synchronisation vectors of all pNets 
%%   included in  $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in K}\myrangle$ 
%%   are finite, and if
%% $\forall i \in L.\, finite{(states(pLTS_i))} \text{ and } pLTS_i$
%% has a finite number of state variables, then Algorithm 1 terminates
%% and produces an open automaton 
%% $\mathcal{T}$ with finitely many states and transitions.

% Given an open pNet $\mylangle \overline{\pNet},\overline{S}, \symb{SV}_k^{k\in
%     K}\myrangle$ with leaves $pLTS_i^{i\in L}$ and holes $Hole_j^{j\in
%   J}$,
% build its semantics as in algorithm 1.

% We have:

% $$ finite{(L)} \land finite{(J)} \land \forall i \in L finite{(\{s_i\})}
%   \to finite{(\mathcal{T})}$$
% \end{theorem}

\TODO{Eric: full result of the tool execution on the running example}

@@

The ``saturation'' step in the construction of the weak relation will
pick all $\tau^*$ and $\tau^*;a$ sequences from the OA, and build a weak
OT from each sequence. Here is one example:

\noindent
$\ostate{1000}\OTarrow{m\_rec(m,x), \phi_1 = [x=s_1\_b \land
    m=s_1\_m]}\ostate{2100}$\\
$\qquad\ostate{2100}$ variables: $\psi_1 = \{m_1\_m:=m; m_1\_x:=x; s_2\_b:=s_1\_b\}$;\\
\\
$\ostate{2100}\OTarrow{\tau <m\_send(m_1\_m,m_1\_x), r\_rec(?m',?b)>,
  \phi_2 = [r_0\_b=b \land m'=m_1\_m \land b=m_1\_x]}\ostate{2001}$\\
$\qquad\ostate{2001}$ variables: $\psi_2 = \{r_1\_b:=r_0\_b; r_1\_m:=m;
; s_2\_b:=s_1\_b\}$;\\
\\
$\ostate{2001}\OTarrow{v,
  \phi_3 = [v=out(r_1\_m)]}\ostate{2002}$\\
$\qquad\ostate{2002}$ variables: $\psi_3 = \{r_2\_b:=r_1\_b; s_2\_b:=s_1\_b\}$;\\

Resulting Weak Transition:

\noindent
$\ostate{1000}\OTWeakarrow{v,
  \phi = \phi_1\land\phi_2\psi_1\land\phi_3\psi_2\psi_1}\ostate{2002}$\\
\\
$\phi=[x=s_1\_b \land m=s_1\_m\\
  \land r_0\_b=b \land m'=m_1\_m \land b=m_1\_x\\
  \land v=out(r_1\_m)]\\
and \\
\psi= \{r_2\_b:=r_0\_b\}$

Note that in $\phi$ the application of assignments
$\psi_1,\psi_2,\psi_3$ have already ``eliminated'' the intermediate
state variables
$m_1\_m, b=m_1\_x, r_1\_m$, but not the intermediate input variables
$m,m',x,b$. But now $\phi$ includes enough information to eliminate them, giving
a final $\phi'= [v=out(s_1\_m)\land r_0\_b=s_1\_b]$

>>>>

\TODO{Eric: refine the algorithm description - \ERIC{Done, can you check ?}}

@@

2) While there are unexplored states, pick one state from the
unexplored set and add it to the explored set.

2a- From this state
build all possible deduction trees by application of the structural
rules Tr1 and Tr2, using all applicable combinations
of synchronisation vectors.
In practice, the algorithm does not explicitely construct the
deduction tree, but for each pNet node, builds all possible
combinations of the OTs of its subnets, and filter these combinations
by matching with all synchronisation vectors.

2b- For each OT built, add the transition in the
outgoing transitions of the current state, and add the
resulting state in the unexplored set if it is not already there.

3) The predicate is submitted to Z3 for checking satisfiability. If it
is NOT satisfiable, the resulting OT is discarded. This will minimize
the number of resulting transitions, and by reachability analysis,
reduce the resulting open automaton.

\end{algorithm}

Given an open-pNet
with finite synchronisation sets, finitely many leaves and
holes, and each pLTS at leaves having a finite number of states and
(symbolic) transitions, has a finite automaton (see \ref{henrio:Forte2016}).

@@

While building the open transitions, many new variable names are
created, as fresh local variables of synchronisation vectors, and
fresh result variables of intermediate OTs. While these are required
for the semantic construction, they are not significant in the
resulting open automaton. In fact they carry information on the
internal structure of the pNet system, and this information should
\_not\_ be visible when looking at the system as a black box, and in
particular when comparing its behaviour with another system, e.g. by
bisimulation.

In the resulting predicate the only significant variables are :\\
- the input variables of pLTS transitions \TODO{heu, non, meme pas, si
  ce sont des comm internes!}\\
- the actions of holes\\
- the result action at toplevel.

So we transform the predicates of the resulting OTs, eliminating all
intermediate variables. \TODO{Explain why this should work !! Maybe
  need to be more precise on unification in the matching...}

\TODO{Eric: full (???) result of the tool execution on the running example}

@@

The ``saturation'' step in the construction of the weak relation will
pick all $\tau^*$ and $\tau^*;a;\tau^*$ sequences from the OA, and build a weak
OT from each sequence. Here is one example:

\noindent
$\ostate{1000}\OTarrow{m\_rec(m,x), \phi_1 = [x=s_1\_b \land
    m=s_1\_m]}\ostate{2100}$\\
and $\qquad\ostate{2100}$ variable assignments are: $\psi_1 = \{m_1\_m:=m; m_1\_x:=x; s_2\_b:=s_1\_b\}$;\\
\\
$\ostate{2100}\OTarrow{\tau <m\_send(m_1\_m,m_1\_x), r\_rec(?m',?b)>,
  \phi_2 = [r_0\_b=b \land m'=m_1\_m \land b=m_1\_x]}\ostate{2001}$\\
and $\qquad\ostate{2001}$ assignments: $\psi_2 = \{r_1\_b:=r_0\_b; r_1\_m:=m;
; s_2\_b:=s_1\_b\}$;\\
\\
$\ostate{2001}\OTarrow{v,
  \phi_3 = [v=out(r_1\_m)]}\ostate{2002}$\\
and $\qquad\ostate{2002}$ variablesassignments: $\psi_3 = \{r_2\_b:=r_1\_b; s_2\_b:=s_1\_b\}$;\\

The resulting Weak Transition has a predicate $\phi =
\phi_1\land\phi_2\psi_1\land\phi_3\psi_2\psi_1 = \phi=[x=s_1\_b \land m=s_1\_m
  \land r_0\_b=b \land m'=m_1\_m \land b=m_1\_x
  \land v=out(r_1\_m)]$
and assignment
$\psi= \{r_2\_b:=r_0\_b\}$.

We can see here that in $\phi$ the application of assignments
$\psi_1,\psi_2,\psi_3$ have already ``eliminated'' the intermediate
state variables
$m_1\_m, b=m_1\_x, r_1\_m$, and not the intermediate input variables
$m,m',x,b$. But now $\phi$ includes enough information to eliminate them, giving
a final $\phi'= [v=out(s_1\_m)\land r_0\_b=s_1\_b]$, so the resulting
WOT is:

 \noindent
 $$\ostate{1000}\OTWeakarrow{v,
   [v=out(s_1\_m)\land r_0\_b=s_1\_b],\{r_2\_b:=r_0\_b\}} \ostate{2002}$$
%% \\
%% $\phi=[x=s_1\_b \land m=s_1\_m\\
%%   \land r_0\_b=b \land m'=m_1\_m \land b=m_1\_x\\
%%   \land v=out(r_1\_m)]\\
%% and \\
%% \psi= \{r_2\_b:=r_0\_b\}$

%% Note that
